A Systematic Literature Review on Extensions to Role Based Access Control

### 2013-04-24 paper/sections/analysis.tex
Adding some edits and another analysis and discussion section.

Core RBAC model provides an abstraction of authorization model based on user role assignment (UA) and the role
permission assignment (PA). Since this simple model may not provide a fine-grained access control for sophisticated
security mechanism, a variety of extended RBAC models have been proposed over the years to meet
security requirements. When administrators design a model, it is important to capture an important abstraction to help the model to be enforced in a system.

@@

The core RBAC standard represents four elements such as roles, sessions, users, and permissions, and their relations. While this model is considered fundamental in any RBAC systems, this model can be extended to be applicable for emerging applications such as healthcare and mobile devices. Such extended RBAC models are formally presented to describe its behaviors. In the selected papers, researchers illustrate how an extended model incorporates new element based on the correct RBAC and describe their relations with pre-defined elements in the core RBAC.

@@

RQ4: How are extension to RBAC evaluated theoeetically and in practice?

@@

Providing evaluation of a proposed model is a key component in establishing the models validity.  The papers were examined
for evidence of evaluations ranging from performance to mathematical accuracy to application to real world scenarios.  Further,
for each proposed model, the reviewers looked for evidence that the authors made comparisons between their own model and the base
model as they pertained to claims made by the authors of why their model is needed.  The quantifiable evaluations looked for were:

@@

Based on the diverse evaluation criteria, 12 models presented no evidence of an evaluation.  8 models presented example scenarios
and how application of their model would apply and resolve the situation.  6 of the models provided some form of performance
or complexity analysis of their model.  This included graphs of the model's time to determine authorization as the number of entities
grew, and the size of the role space for the extension model compared to standard RBAC. 4 models provided mathematical descriptions
and analysis as a way to provide evaluation in the form of completeness.

The most widely used evaluation method was providing sample scenarios with accompanying workflows of how the extension model
would tackle those scenarios.  Much is left to the reader to assume of these types of evaluations, as the authors do not explicitly state
or show how the standard model is deficient in tackling said scenarios.

@@

@@

Mobile computing has seen a dramatic increase in the number of available devices, operating systems and applications since 1997 when the first smart phone
was introduced. 
The domain analysis results produced 5 papers that targeted extensions that are designed to address the requirements of mobile computing.
For a domain that was roots in personal and enterprise computing, protecting the data of both through access controls is paramount given their ubiquity.

>>>>

The RBAC standard model provides an abstraction of authorization based on user role assignment (UA) and role
permission assignment (PA). Since this simple model may not provide a fine-grained access control for sophisticated
security mechanism, a variety of extended RBAC models have been proposed over the years to meet
security requirements. When administrators design a model, it is important to capture an important abstraction to help the model to be enforced in a system.

@@

The RBAC standard model presents four entities: roles, sessions, users, and permissions, and their relations. While this model is considered fundamental in any RBAC systems, this model can be extended to be applicable for emerging applications such as healthcare and mobile devices. Such extended RBAC models are formally presented to describe its behaviors. In the selected papers, researchers illustrate how an extended model incorporates new element based on the correct RBAC and describe their relations with pre-defined elements in the core RBAC.

@@

RQ4: How are extension to RBAC evaluated theoretically and in practice?

@@

The 28 primary sources were examined for evidence that evaluations of the proposed model were presented by the model authors.
Further, we identified a set of evaluation types found within the each of the primary sources and provide below a list of the 
evaluation types and which primary sources provided which type.  In some cases a single primary source provided mutliple evaluation
types.

@@

Based on the diverse evaluation criteria, 12 models presented no evidence of an evaluation. Eight models presented example scenarios
and how application of their model would apply and resolve the situation.  Six of the models provided some form of performance
or complexity analysis of their model.  This included graphs of the model's time to determine authorization as the number of entities
grew, and the size of the role space for the extension model compared to standard RBAC. Four models provided mathematical descriptions
and analysis as a way to provide evaluation in the form of completeness. 
The most widely used evaluation method was providing sample scenarios with accompanying workflows of how the extension model
would tackle those scenarios. Much is left to the reader to assume of these types of evaluations, as the authors do not explicitly state
or show how the standard model is deficient in tackling said scenarios.

@@

When proposing an access control model, providing an evaluation of the model is a key component in establishing the validity of the model. 
Further, in the case of extensions to the RBAC standard model, the model should be accompanied by validation of the model as a stand-alone
access control model and in comparison to the model upon which the enhancements are being made. The results show that robust evaluations of
extenion models are lacking.

The primary source of validation a developer or practioner may encounter is a qualitative discussion of real-world
scenarios and how the proposed model can tackle those situations. In rare cases, five primary sources, the model 
authors provide some discussion of how the RBAC standard model is deficient in tackling the scenario. In rarer cases, 
one paper from our results, a case study is performed to examine how the proposed model works in practice. Discussions
of how an extension model handles a real-world scenario provides developers and practioners ancedotal evidence at best
for what types of situations the proposed model could handle. Further, by not providing a comparison to the RBAC standard
model, developers may be left implementing a more complex model to address their requirements when the standard model would
have sufficed.  Further, given the nature of access control models as grounded in application to enterprise implementations, 
case studies of a model in action provide developers with evidence that the model works as intended when applied.

When looking for an enterprise ready access control mechanism, developers must balance usability with security. 
Two of the primary sources examined provided time-based performance analysis of their extension model compared
to the RBAC standard model. This inclusion of time analysis provides some assurances to developers that any non-functional
requirements surrounding time to compute authorizations compete or beat the standard model. Further, four of the models provided
some form of complexity of their model. This complexity analysis plays a key role in the management of the access control mechanism
over the course of the models implementation lifetime. As the number of roles, users and additional entities grows, developers will 
need to ensure non-functional requirements are met that deal with the ability for a system administrator to effectively manage these
entities.

@@

Mobile computing has seen a dramatic increase in the number of available devices, operating systems and applications since 1997 when the first smart phone
was introduced. 
The domain analysis results produced 5 papers that targeted extensions that are designed to address the requirements of mobile computing.
For a domain that has roots in personal and enterprise computing, protecting the data of both through access controls is paramount given their ubiquity.

### 2013-04-24 paper/sections/core_rbac.tex
Adding some edits and another analysis and discussion section.

RBAC provides effective and efficient permissions management for operations, especially when sharing resources in an organizations.
Prior to the creation of the NIST RBAC standard, no general agreement on the definition 
of RBAC existed amongst practioners or within the research community. 
Without a unified definition of RBAC, software developers describe similar concepts and features of RBAC models using different terminology, which was shown to slow implementation of RBAC.  
Moreover, in cases where organizations are concerned with adopting RBAC,
evaluation and comparison of RBAC technologies developed by different vendors was difficult.
NIST, in collaboration with industry and academics, worked on defining a set of consensus RBAC concepts and terminology and proposed a standard for RBAC that addressed these cost and interoperability issues by developing a common definition that can be used across different vendors.

>>>>

RBAC provides effective and efficient permissions management for operations, especially when sharing resources withinin an organization.
Prior to the creation of the NIST RBAC standard, no general agreement on the definition 
of RBAC existed amongst practioners or within the research community. 
Without a unified definition of RBAC, software developers described similar concepts and features of RBAC models using different terminology. 
This lack of consistent terminology was shown to slow the implementation of RBAC \cite{o20102010}.  
Moreover, in cases where organizations were concerned with adopting RBAC,
evaluation and comparison of RBAC technologies developed by different vendors was difficult.
NIST, in collaboration with industry and academics, worked on defining a set of consensus RBAC concepts and terminology and proposed a standard for 
RBAC that addressed these cost and interoperability issues by developing a common definition that can be used across different vendors.

### 2013-04-24 paper/sections/intro.tex
Adding some edits and another analysis and discussion section.

Access control models controls which principal can access which resources in what manner.
Role based access control (RBAC) is one of access control models, which is used especially for maintaining and managing an organization's access control based on roles instead of individual users. The RBAC model has become popular since the National Institute for Standards and Technology (NIST) first proposed the RBAC standard in the 1990s. The RBAC model is widely used for protection of security and privacy of various applications including web services, database applications, and healthcare applications. 
In the RBAC model, roles represent a set of permissions needed to perform a particular job function within an organization.  
Any number of users who are involved in that specific job function within the organization can then be assigned to the role to inherit the required access instead of assigning individual permissions to individual users. 
As the number of permissions and users scales upward, the ability to logically group both users and permissions into smaller numbers of entities becomes paramount for managing access control as an organization grows and evolves. The RBAC model has advantage for managing access control. For example, if a user requires access to resources associated with the manager role within a given organization, security policy administrators need only associate the user with the manager role.

@@

\textit{The goal of our work is to aid practioners and researchers in choosing an extension to the RBAC standard model, and in understanding
how extensions to RBAC are evaluated by providing an assessment of the state of extensions to the RBAC standard model through: establishment of a 
set of extension categories, an examination of the state of the art in evaluations of extension models and a breakdown of the reasoning and motivations
that have led to extension of the RBAC standard.} With respect to our goal, the authors addressed the following research questions:

>>>>

Role based access control (RBAC) is used for maintaining and managing an organization's access control based on assigning permissions to roles instead of assigning the permissions directly to individual users. The RBAC model has become popular since the National Institute for Standards and Technology (NIST) first proposed the RBAC standard in the 1990s. The RBAC model is widely used for protection of security and privacy of various applications including web services, database applications, and healthcare applications. 
In the RBAC model, roles represent a set of permissions needed to perform a particular job function within an organization.  
Any number of users who are involved in that specific job function within the organization can then be assigned to the role to inherit the required access instead of assigning individual permissions to individual users. 
As the number of permissions and users scales upward, the ability to logically group both users and permissions into smaller numbers of entities becomes paramount for managing access control as an organization grows and evolves. The RBAC model has advantage for managing access control. For example, if a user requires access to resources associated with the manager role within a given organization, security policy administrators need only associate the user with the manager role.

@@

\textit{The goal of our work is to aid practitioners and researchers in choosing an extension to the RBAC standard model, and in understanding
how extensions to RBAC are evaluated by providing an assessment of the state of extensions to the RBAC standard model. We accomplished this through: 
establishment of a set of extension categories, an examination of the state of the art in evaluations of extension models and a breakdown of the 
motivations that have led to extension of the RBAC standard model.} With respect to our goal, the authors addressed the following research questions:
